---
title: "Spatial_dataset_preprocessing"
author: "Dillon Corvino"
date: "10/07/2024"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: united
    highlight: tango
    df_print: paged
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

Built with R version `r getRversion()`

## Setup {.tabset}

### Dataset information
```{r Dataset_Info}

# Human 
# RCC - subjected to targeted spatial RNA analysis 
# dataset provided by Marieta Toma 


```

### Environment
```{r setup}

knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,  
  eval = TRUE, 
  tidy = TRUE
)

knitr::opts_knit$set(
  root.dir = "../"
)


# Environment Set up
rm(list = ls()) #Clean workspace
cat("\014")     #Clean Console
gc() # Free memory

# pipeline variables
long.compute <- TRUE
do.plot <- TRUE

# Establish working directory
library("rstudioapi")

# Set working directory to source file location
setwd(dirname(getActiveDocumentContext()$path))
setwd("..")

# Create output directories & load custom functions & colour scheme
source("scripts/Setup.R", local = knitr::knit_global())


ggplot2_switch_version(desired_version = "3.4.4")


# load packages
library("Seurat")
library("SeuratObject")
library("SeuratData")
library("Azimuth")
library("SeuratWrappers")
library("dplyr")
library("scCustomize")



```


## Pre-processing {.tabset}


### Load dataset
```{r load_dataset}

if(long.compute){
  input.path <- "/Users/ieo/Documents/Sequence_alignment/Marieta_toma_Spatial_RCC_dataset/"
  region.1.path <- "output-XETG00229__0019014__Region_1__20240131__145505/"
  region.2.path <- "output-XETG00229__0019014__Region_2__20240131__145505/"
  
  region.1 <- LoadXenium(data.dir = paste0(input.path, region.1.path))
  region.2 <- LoadXenium(data.dir = paste0(input.path, region.2.path))
  
  
  
  # remove cells with 0 counts
  region.1 <- subset(region.1, subset = nCount_Xenium > 0)
    region.2 <- subset(region.2, subset = nCount_Xenium > 0)

  rownames(region.1)
  
  ImageDimPlot(region.2, fov = "fov", molecules = c("NKG7", "CD8A", "FOXP3", "CD19"), nmols = 500, axes = T)
  
  VlnPlot(region.1, features = c("nFeature_Xenium", "nCount_Xenium"), ncol = 2, pt.size = 0)

  ImageFeaturePlot(region.2, features = c("NKG7", "CD8A", "FOXP3", "CD19"), max.cutoff = c(25,
    35, 12, 10), size = 0.75, cols = c("white", "red"))
  
  
  
  region.1 <- SCTransform(region.1, assay = "Xenium")
  region.1 <- RunPCA(region.1, npcs = 30, features = rownames(region.1))
  region.1 <- RunUMAP(region.1, dims = 1:30)
  region.1 <- FindNeighbors(region.1, reduction = "pca", dims = 1:30)
  region.1 <- FindClusters(region.1, resolution = 0.3)
  
  DimPlot(region.1)
  FeaturePlot_scCustom(region.1, c("CD8A", "CD4"))
  
    VlnPlot(region.1, features = c("CD8A", "CD4"), ncol = 2, pt.size = 0, group.by = "orig.ident")


      
      
      ImageDimPlot(region.2, cols = "polychrome", size = 0.75)
  
#devtools::install_github("dmcable/spacexr", build_vignettes = FALSE)
  
library(spacexr)

query.counts <- GetAssayData(region.1, assay = "Xenium", layer = "counts")
coords <- GetTissueCoordinates(region.1, which = "centroids")
rownames(coords) <- coords$cell
coords$cell <- NULL
query <- SpatialRNA(coords, query.counts, colSums(query.counts))




}

```



## QC and Filter {.tabset}

### QC & filter

```{r QC_and_filter}

if(long.compute){
  
  
  for(i in seq_along(sequencing.runs)){
    
    
    run.id <- unique(HL.seurat.list[[i]]$orig.ident)
    
    
    # Create output dir
    
    output.dir <- paste0("results/QC/QC_", run.id)
    
    if(!dir.exists(output.dir)){
      dir.create(output.dir, 
                 recursive = T)
      
    }
    
    
    # set idents to sampleID
    Idents(HL.seurat.list[[i]]) <- HL.seurat.list[[i]]$orig.ident
    
    # set default assay 
    DefaultAssay(HL.seurat.list[[i]]) <- "RNA"
    
    # Add mitochondrial and ribosomal gene percentages
    
    HL.seurat.list[[i]] <- Add_Mito_Ribo(object = HL.seurat.list[[i]], species = "Human", overwrite = TRUE) 
    
    # Add cell complexity/novelty QC metric
    
    HL.seurat.list[[i]] <- Add_Cell_Complexity(object = HL.seurat.list[[i]])
    
    # Setting cutoff values 
    feature.low <- 650
    feature.high <- 5500
    umi.low <- 1200
    umi.high <- 45000
    mito.high <- 10
    ribo.high <- 6
    complexity.high <- 0.8
    
    
    ########################
    # QC plots
    ########################
    if(do.plot){
      # nFeatures
      print(
        QC_Plots_Genes(seurat_object = HL.seurat.list[[i]], 
                       low_cutoff = feature.low, 
                       high_cutoff = feature.high,
                       pt.size = 0.05) #how many genes per cell? See where to set the cutoffs later
      )
      
      dev.copy(pdf, paste0(output.dir, "_genes.pdf"))
      dev.off()
      
      
      # UMI
      print(
        QC_Plots_UMIs(seurat_object = HL.seurat.list[[i]], 
                      low_cutoff = umi.low, 
                      high_cutoff = umi.high, 
                      y_axis_log = T,
                      pt.size = 0.05)
      )
      
      dev.copy(pdf, paste0(output.dir, "_UMI.pdf"))
      dev.off()
      
      
      # Mito percentage
      print(
        QC_Plots_Mito(seurat_object = HL.seurat.list[[i]], 
                      high_cutoff = mito.high,
                      pt.size = 0.05)
      )
      
      dev.copy(pdf, paste0(output.dir, "_mito.pdf"))
      dev.off()
      
      
      # Ribo percentage
      print(
        QC_Plots_Feature(seurat_object = HL.seurat.list[[i]],
                         feature = "percent_ribo",
                         high_cutoff = ribo.high,
                         pt.size = 0.05)
      )
      
      dev.copy(pdf, paste0(output.dir, "_Ribo.pdf"))
      dev.off()
      
      
      # Complexity
      print(
        QC_Plots_Complexity(seurat_object = HL.seurat.list[[i]], 
                            high_cutoff = complexity.high,
                            pt.size = 0.05)
      ) 
      
      dev.copy(pdf, paste0(output.dir, "_complexity.pdf"))
      dev.off()
      
    }
    
    # Filter dataset
    
    HL.seurat.list[[i]] <- subset(HL.seurat.list[[i]], 
                                  subset = nFeature_RNA > feature.low & 
                                    nFeature_RNA < feature.high & 
                                    percent_mito < mito.high & 
                                    percent_ribo > ribo.high & 
                                    log10GenesPerUMI > complexity.high)
    
    if(do.plot){
      # nFeatures filtered matrix
      print(
        QC_Plots_Genes(seurat_object = HL.seurat.list[[i]], 
                       low_cutoff = feature.low, 
                       high_cutoff = feature.high,
                       pt.size = 0.05) #how many genes per cell? See where to set the cutoffs later
      )
      
      dev.copy(pdf, paste0(output.dir, "_genes_filtered.pdf"))
      dev.off()
      
      
      # UMI filtered matrix
      print(
        QC_Plots_UMIs(seurat_object = HL.seurat.list[[i]], 
                      low_cutoff = umi.low, 
                      high_cutoff = umi.high, 
                      y_axis_log = T,
                      pt.size = 0.05)
      )
      
      dev.copy(pdf, paste0(output.dir, "_UMI_filtered.pdf"))
      dev.off()
      
      # Ribo filtered percentage
      print(
        QC_Plots_Feature(seurat_object = HL.seurat.list[[i]],
                         feature = "percent_ribo",
                         high_cutoff = ribo.high,
                         pt.size = 0.05)
      )
      
      dev.copy(pdf, paste0(output.dir, "_Ribo_filtered.pdf"))
      dev.off()
      
      
      
      ##########################################
      # check ADT and hashtag datasets
      ##########################################
      
      head(HL.seurat.list[[i]]@meta.data)
      
      # nCount ADT dataset
      print(
        VlnPlot(HL.seurat.list[[i]],
                features = "nCount_ADT",
                pt.size = 0.05) + NoLegend()
      )
      
      dev.copy(pdf, paste0(output.dir, "_nCount_ADT.pdf"))
      dev.off()
      
      # nFeature ADT dataset
      print(
        VlnPlot(HL.seurat.list[[i]], 
                features = "nFeature_ADT",
                pt.size = 0.05) + NoLegend()
      )
      dev.copy(pdf, paste0(output.dir, "_nFeature_ADT.pdf"))
      dev.off()
      
      # nCount hashtag dataset
      print(
        VlnPlot(HL.seurat.list[[i]], 
                features = "nCount_Hashtag", 
                pt.size = 0.05) + NoLegend()
      )
      
      dev.copy(pdf, paste0(output.dir, "_nCount_hashtag.pdf"))
      dev.off()
      
      # nfeature hashtag dataset
      print(
        VlnPlot(HL.seurat.list[[i]],
                features = "nFeature_Hashtag", 
                pt.size = 0.05) + NoLegend()
      )
      
      dev.copy(pdf, paste0(output.dir, "_nfeature_hashtag.pdf"))
      dev.off()
      
    }
    
  }
  
}

```



### Save seurat object
```{r save_seurat_object}

if(long.compute){
  
  # export and save seurat object
  # SaveH5Seurat(HL.seurat,
  #              filename = "saves/HL_seurat.h5Seurat",
  #              overwrite = TRUE, 
  #              verbose = TRUE)
  # 
  
  SaveSeuratRds(HL.seurat, file = "saves/Preprocess_HL_seurat.rds")
  
}

```

## Adding TCR data {.tabset}

### Create T-cell only seurat object
```{r Tcell_seurat_object}

# set Idents
Idents(HL.seurat) <- HL.seurat@meta.data$anno.l1
DefaultAssay(HL.seurat) <- "SCT"

# subset to just T-cell populations, a handful of TCRs are found in monocytes, B-cells, NK cells but these will be ignored
Tcell.seurat <- subset(HL.seurat, idents = c("CD4_cells", "CD8_cells"))
Tcell.seurat@meta.data$anno.l1 <- droplevels(Tcell.seurat@meta.data$anno.l1)
Tcell.seurat@meta.data$anno.l2 <- droplevels(Tcell.seurat@meta.data$anno.l2)



# note that did look if TCR map to MAIT/ILC populations and found 
# out of this many cells
# ILCs = 169
# MAITs = 1,554
# Tregs = 147

# TCR was recovered for this many cells
# MAITs = 77 cells
# ILCs = 2
# Tregs = 15

# Therefore represents approx. 5% of MAIT cells have a TCR
# therefore dont include MAIT or ILC populations 

```


### Format data with scRepertoire
```{r Format_data_with_scRep}

#################################
# Read in and format data
#################################

# install packages
#BiocManager::install("scRepertoire")


# Load library
library("scRepertoire")

# read in contig files
# note that sequence run 1 doesnt have vdj sequencing
run.2.contig <- loadContigs("data/Run_2_allsamples/vdj_t/", format = "10X")
run.3.contig <- loadContigs("data/Run_3_allsamples/vdj_t/", format = "10X")
run.4.contig <- loadContigs("data/Run_4_allsamples/vdj_t/", format = "10X")




# create list of contigs
contig.list <- list(run.2.contig, run.3.contig,
                    run.4.contig)

# Combine contig list
combined.contig <- combineTCR(contig.list, 
                              samples = c("Run_2", "Run_3", "Run_4"), 
                              #ID = c("Healthy", "Patient", "Patient"), 
                              removeNA = TRUE,
                              removeMulti = FALSE, 
                              filterMulti = TRUE) # take the top 2 expressed chains for cells with multiple chains 

# on average, filteringMulti results in approx. 200 more cells for each sample than outright removing cells
head(combined.contig)
colnames(HL.seurat)


# dataset summary
table(Tcell.seurat@meta.data$anno.l1) 
table(Tcell.seurat@meta.data$anno.l2) 
# CD4_cells = 21,886
# CD8_cells = 5,857
# Tregs = 147 cells


# Combine seurat object with contig info combineExpression
HL.seurat.tcr <- combineExpression(combined.contig,
                                   Tcell.seurat,
                                   cloneCall = "aa",
                                   chain = "both",
                                   proportion = TRUE,
                                   cloneSize = c(Rare = 1e-04, 
                                                 Small = 0.001,
                                                 Medium = 0.01, 
                                                 Large = 0.1, 
                                                 Hyperexpanded = 1),
                                   group.by = "none", # Alternative frequencies are calculated in next code Chunk
                                   filterNA = TRUE) # should seurat object be subsetted to remove NA values

# Summarise tcrs captured

table(HL.seurat.tcr@meta.data$anno.l1) 
table(HL.seurat.tcr@meta.data$anno.l2) 
# CD4_cells = 2,790
# CD8_cells = 571
# Tregs = 15 cells

# this is equal to
(2790/21886)*100 # ~13 percent of CD4s have a TCR found
(571/5857)*100 # ~10 percent of CD8s have a TCR found
(15/147)*100 # ~10 percent of Tregs have a TCR found


lapply(combined.contig, dim)
# run 2 = 2,674 
# run 3 = 1,792 
# run 4 = 601
# total = 5067
(3376/5067)*100 # 66% of sequenced TCRs are recovered and applied to dataset



# Write out data for future use in scRep package
saveRDS(combined.contig, "saves/combined_contig.rds")


# remove unneeded objects
rm(contig.list, run.2.contig, run.3.contig, run.4.contig)

```

### Split metadata CTgene into TRAV/J/C and TRBV/J/C gene calls
```{r split_CTgene_metadata_info}

library(stringr)


# Split the CTgene string at "." 
temp.list <- str_split(HL.seurat.tcr@meta.data$CTgene, "\\.")
temp.list


# First and second elements of the list are V and J calls
HL.seurat.tcr@meta.data$TRAV <- sapply(temp.list, "[[", 1)
HL.seurat.tcr@meta.data$TRAJ <- sapply(temp.list, "[[", 2)

# Third element is the TRAC and TRBV combined - therefore split this by "_"
temp.list.2 <- sapply(temp.list, "[[", 3)
temp.list.2 <- str_split(temp.list.2, "_")

# First element is TRAC and second is TRBV gene call
HL.seurat.tcr@meta.data$TRAC <- sapply(temp.list.2, "[[", 1)
HL.seurat.tcr@meta.data$TRBV <- sapply(temp.list.2, "[[", 2)

# 4th and 5th elements of original list are TRBJ and TRBC calls respectively
HL.seurat.tcr@meta.data$TRBJ <- sapply(temp.list, "[[", 4)
HL.seurat.tcr@meta.data$TRBC <- sapply(temp.list, "[[", 6)

# remove unneeded variables 
rm(temp.list, 
   temp.list.2)

```

### Count clonotype frequencies and normalise
```{r Count_and_normalise_clonotype_frequencies}

####################################################################################
# Need to count clonotype occurrence per donor and then normalise this value 
####################################################################################

# create temp meta.data df and append cell_ID as column (lost in dplyr workflow)
meta.data.temp <- HL.seurat.tcr@meta.data
meta.data.temp$cell_ID <- rownames(HL.seurat.tcr@meta.data)


# Get total number of cells in each donor that has a TCR
Donor.cell.n <- table(meta.data.temp$DonorID)
total.cell.n <- sum(Donor.cell.n) # 3,376 cells total


####################################################################################
# Count occurence of clonotypes across entire dataset 
####################################################################################
# Also: Normalise n() by # of cells within dataset
meta.data.temp <- meta.data.temp %>%
  dplyr::group_by(CTaa) %>%
  dplyr::mutate(Total_clonotype_n = n()) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(Total_clonotype_freq = (Total_clonotype_n/total.cell.n)*100) %>%
  dplyr::ungroup()

####################################################################################
# Count occurence of clonotypes across entire condition 
####################################################################################
# Also: Normalise n() by # of cells within condition
meta.data.temp <- meta.data.temp %>%
  dplyr::group_by(CTaa, DonorID) %>%
  dplyr::mutate(Clone_Donorwise_n = n()) %>%
  dplyr::group_by(DonorID) %>%
  dplyr::mutate(Clone_Donorwise_freq = case_when(DonorID == "Unknown" ~ (Clone_Donorwise_n/Donor.cell.n["Unknown"])*100,
                                                 DonorID == "HD_03" ~ (Clone_Donorwise_n/Donor.cell.n["HD_03"])*100, 
                                                 DonorID == "HD_04" ~ (Clone_Donorwise_n/Donor.cell.n["HD_04"])*100,
                                                 DonorID == "Pt_01" ~ (Clone_Donorwise_n/Donor.cell.n["Pt_01"])*100,
                                                 DonorID == "Pt_02" ~ (Clone_Donorwise_n/Donor.cell.n["Pt_02"])*100,
                                                 DonorID == "Pt_03" ~ (Clone_Donorwise_n/Donor.cell.n["Pt_03"])*100,
                                                 DonorID == "Pt_04" ~ (Clone_Donorwise_n/Donor.cell.n["Pt_04"])*100,
                                                 DonorID == "Pt_05" ~ (Clone_Donorwise_n/Donor.cell.n["Pt_05"])*100,
                                                 DonorID == "Pt_06" ~ (Clone_Donorwise_n/Donor.cell.n["Pt_06"])*100,
                                                 DonorID == "Pt_07" ~ (Clone_Donorwise_n/Donor.cell.n["Pt_07"])*100,
                                                 DonorID == "Pt_08" ~ (Clone_Donorwise_n/Donor.cell.n["Pt_08"])*100,
                                                 DonorID == "Pt_09" ~ (Clone_Donorwise_n/Donor.cell.n["Pt_09"])*100,
                                                 DonorID == "Pt_10" ~ (Clone_Donorwise_n/Donor.cell.n["Pt_10"])*100,
                                                 DonorID == "Pt_11" ~ (Clone_Donorwise_n/Donor.cell.n["Pt_11"])*100,
                                                 DonorID == "Pt_12" ~ (Clone_Donorwise_n/Donor.cell.n["Pt_12"])*100)) %>%
  dplyr::ungroup()


######################################
# Add new metadata to seurat object
######################################

#########
# Total 
#########

# Add total clonotype n value
HL.seurat.tcr <- AddMetaData(object = HL.seurat.tcr, 
                             col.name = "Total_clonotype_n",
                             metadata = meta.data.temp$Total_clonotype_n)

# Add total clonotype freq
HL.seurat.tcr <- AddMetaData(object = HL.seurat.tcr, 
                             col.name = "Total_clonotype_freq",
                             metadata = meta.data.temp$Total_clonotype_freq)

##################
# Donor-wise 
##################

# Add clone Donor-wise n value
HL.seurat.tcr <- AddMetaData(object = HL.seurat.tcr, 
                             col.name = "Clone_Donorwise_n",
                             metadata = meta.data.temp$Clone_Donorwise_n)

# Add Clone Donor-wise freq
HL.seurat.tcr <- AddMetaData(object = HL.seurat.tcr, 
                             col.name = "Clone_Donorwise_freq",
                             metadata = meta.data.temp$Clone_Donorwise_freq)



rm(meta.data.temp)

```


### Annotate cells as rare, small, hyperexpanded etc and clean metadata
```{r annotate_clonotype}


# clonotype calculation done by scRep doesnt look right, use my own abundance calculation and re-annotate clonotype


# Firstly, remove freq data and clonotype call, and indicator generated by scRep
head(HL.seurat.tcr@meta.data)

HL.seurat.tcr@meta.data$clonalFrequency <- NULL
HL.seurat.tcr@meta.data$cloneSize <- NULL
HL.seurat.tcr@meta.data$clonalProportion <- NULL
head(HL.seurat.tcr@meta.data)




# get clonotype abundance stats


summary(HL.seurat.tcr@meta.data$Total_clonotype_n) # 22 = max
summary(HL.seurat.tcr@meta.data$Clone_Donorwise_n) # 19 = max

clonotype.vals <- HL.seurat.tcr@meta.data$Clone_Donorwise_n
cloneType <- HL.seurat.tcr@meta.data$Clone_Donorwise_n * NA

single = 1
small = 5
medium = 10
large = 20
hyperexpanded = 150

cloneType[clonotype.vals == single] <- "Single (1)"
cloneType[clonotype.vals > single & clonotype.vals <= small] <- "Small (1 < X <= 5)"
cloneType[clonotype.vals > small & clonotype.vals <= medium] <- "Medium (5 < X <= 10)"
cloneType[clonotype.vals > medium & clonotype.vals <= large] <- "Large (10 < X <= 20)"
cloneType[clonotype.vals > large & clonotype.vals <= hyperexpanded] <- "Hyperexpanded (20 < X <= 150)"
unique(cloneType)


# Ensure seurat object clonoType var is a factor with correct levels
HL.seurat.tcr@meta.data$Clonetype <- factor(cloneType, 
                                            levels = c("Hyperexpanded (20 < X <= 150)",
                                                       "Large (10 < X <= 20)", 
                                                       "Medium (5 < X <= 10)",
                                                       "Small (1 < X <= 5)", 
                                                       "Single (1)"))

HL.seurat.tcr@meta.data$Clonetype <- droplevels(HL.seurat.tcr@meta.data$Clonetype)

unique(HL.seurat.tcr@meta.data$Clonetype)
```

### Add TCR data to T-cell only seurat
```{r add_TCR_to_Tcell_seurat}


# It was easier to use the filtering option of scRep to remove anything without a TCR, then do filtering such as splitting TCRA and TCRB gene strings
# then add this TCR data back to the Tcell only seurat object
HL.seurat.tcr # 3,376 cells
Tcell.seurat # 27,890 cells

# just select metadata from the tcr object
temp.metadata <- HL.seurat.tcr@meta.data %>%
  dplyr::select(c(CTgene, CTnt, CTaa, CTstrict, 
                  TRAV, TRAJ, TRAC, TRBV, TRBJ, TRBC, 
                  Total_clonotype_n, Total_clonotype_freq, Clone_Donorwise_n, Clone_Donorwise_freq, Clonetype)) %>%
  tibble::rownames_to_column()

# extract Tcell only seurat metadata
Tcell.metadata <- Tcell.seurat@meta.data %>%
  tibble::rownames_to_column()

# merge metadata objects
new.metadata <- merge(Tcell.metadata, temp.metadata, by = "rowname", all = TRUE)

new.metadata <- new.metadata %>%
  tibble::column_to_rownames()

sum(rownames(new.metadata) != rownames(Tcell.seurat@meta.data)) # data is in exact same order

# add the entire metadata table to seurat object
Tcell.seurat@meta.data <- new.metadata

# further annotate for easy downstream plotting and filtering
Tcell.seurat@meta.data <- Tcell.seurat@meta.data %>%
  dplyr::mutate(TCRStatus = ifelse(is.na(CTaa), "NO", "YES"))

# remove unneeded objects
rm(HL.seurat.tcr)
rm(temp.metadata)
rm(new.metadata)
rm(Tcell.metadata)

```


### GEX normalisation & integration Tcell only seurat
```{r GEX_norm_and_integration_Tcell_only}

# GEX normalisation
if(long.compute){
  
  # Note, 
  # I compared CCA integration and rpca integration, looks cca is better fit for this dataset 
  # also see documentation on cca vs rpca integration, cca more aligns with this dataset 
  
  # set default assay 
  DefaultAssay(Tcell.seurat) <- "RNA"
  
  # split object up based on sequencing run / in seurat V5 these are splitting the layers
  #Tcell.seurat[["RNA"]] <- split(Tcell.seurat[["RNA"]],  f = Tcell.seurat$RunID)
  
  
  
  #####################################################
  # Normalise with scTransform and integrate data
  #####################################################
  # note error with more recent version of matrixStats resulting in "useNames = NA is defunct" error. downgrade to matrixStats version 1.1.0 as a workaround
  # https://github.com/satijalab/seurat/issues/7501#issuecomment-1854571904
  #remotes::install_version("matrixStats", version="1.1.0") # restart your session and run previous scripts  
  
  
  # note RNA assay is split into run level layers, below script will apply normalisation etc layer-wise
  
  # Normalise and scale with SCTransform v2
  Tcell.seurat <- SCTransform(Tcell.seurat, 
                              method = "glmGamPoi", 
                              vars.to.regress = "percent_mito", 
                              vst.flavor = "v2",
                              verbose = TRUE) %>%
    RunPCA(reduction.name = "pca_sct") %>%
    RunUMAP(reduction = "pca_sct",
            dims = 1:30, 
            reduction.name = "umap.unintegrated.sct")
  
  DefaultAssay(HL.seurat) <- "SCT"
  
  # integrate layers using cca integration
  Tcell.seurat <- IntegrateLayers(object = Tcell.seurat,
                                  orig.reduction = "pca_sct",
                                  new.reduction = "integrated.cca.sct",
                                  method = CCAIntegration,
                                  normalization.method = "SCT",
                                  verbose = F)
  
  # find clusters and dim reduction using integrated data
  Tcell.seurat <- FindNeighbors(Tcell.seurat,
                                dims = 1:30,
                                reduction = "integrated.cca.sct") %>%
    FindClusters(resolution = 1.5, 
                 cluster.name = "clusters_cca_sct") %>%
    RunUMAP(reduction = "integrated.cca.sct",
            dims = 1:30, 
            reduction.name = "umap.cca.sct")
  
  
  
  
  if(do.plot){
    
    # create output dir
    output.dir <- paste0("results/integration_Tcell_only_seurat/")
    
    if(!dir.exists(output.dir)){
      dir.create(output.dir, 
                 recursive = T)
      
    }
    
    
    unintegrated.reduction <- "umap.unintegrated.sct"
    integrated.reduction <- "umap.cca.sct"
    title.var <- "SCT"
    
    
    
    # evaluate integration and dim reduction
    
    vars.plot <- c("RunID", "predicted.celltype.l1", "DonorID")
    for(i in seq_along(vars.plot)){
      
      
      # remove "." from name
      name.var <- gsub("\\.", "_", vars.plot[i])
      
      before.int <- DimPlot(Tcell.seurat,
                            reduction = unintegrated.reduction, 
                            label = T,
                            group.by = vars.plot[i]) + NoLegend() + ggtitle(paste0("Unintegrated ", title.var, " Normalisation"))
      
      after.int <- DimPlot(Tcell.seurat,
                           reduction = integrated.reduction, 
                           label = T,
                           group.by = vars.plot[i]) + NoLegend() + ggtitle(paste0("CCA Integration using ", title.var, " Normalisation"))
      
      print(before.int | after.int)
      dev.copy(pdf, paste0(output.dir, "UMAP_before_after_integration", name.var, ".pdf"))
      dev.off()
      
      
      
    }
    
    
    
  }
  
}


```

### ADT normalisation & integrate Tcell only seurat
```{r ADT_norm_and_integration_Tcell_only}


if(long.compute){
  
  # create output dir
  output.dir <- paste0("results/integration_Tcell_only_seurat/")
  
  if(!dir.exists(output.dir)){
    dir.create(output.dir, 
               recursive = T)
    
  }
  
  
  # Normalize ADT data,
  DefaultAssay(Tcell.seurat) <- "ADT"
  
  VariableFeatures(Tcell.seurat) <- rownames(Tcell.seurat[["ADT"]])
  Tcell.seurat <- NormalizeData(Tcell.seurat, 
                                normalization.method = "CLR",
                                margin = 2, # normalise accross cells
                                verbose = TRUE) %>%
    ScaleData() %>% 
    RunPCA(reduction.name = 'apca')
  
  
  # Integrate ADT data
  
  # split object up based on sequencing run
  Tcell.seurat[["ADT"]] <- split(Tcell.seurat[["ADT"]],  f = Tcell.seurat$RunID)
  
  # Note: i investigated not integrating, vs rpca vs cca and rpca and cca look quite comparable 
  # there was a decent batch effect evident so integration is needed 
  # for consistency used cca as this was used for gex data also
  
  # integrate layers using cca integration
  Tcell.seurat <- IntegrateLayers(object = Tcell.seurat,
                                  orig.reduction = "apca",
                                  new.reduction = "integrated.cca.adt",
                                  features = rownames(Tcell.seurat[["ADT"]]),
                                  method = CCAIntegration,
                                  verbose = F)
  
  
  
  # Create adt only data umap
  Tcell.seurat <- RunUMAP(Tcell.seurat,
                          reduction = "integrated.cca.adt",
                          dims = 1:30, 
                          reduction.name = "umap.cca.adt")
  
  
  DimPlot(Tcell.seurat,
          reduction = "umap.cca.adt", 
          label = T,
          group.by = "RunID") + 
    NoLegend() + ggtitle(paste0("CCA Integrated adt data only umap"))
  
  dev.copy(pdf, paste0(output.dir, "UMAP_integrated_ADT_data.pdf"))
  dev.off()
  
  DefaultAssay(Tcell.seurat) <- "RNA"
  
}
```


### Weighted Nearest Neighbor Analysis Tcell only seurat
```{r WNNA_Tcell_only}

if(long.compute){
  
  
  DefaultAssay(Tcell.seurat) <- "RNA"
  
  Tcell.seurat <- FindMultiModalNeighbors(Tcell.seurat, 
                                          reduction.list = list("integrated.cca.sct", "integrated.cca.adt"), 
                                          dims.list = list(1:30, 1:47), 
                                          modality.weight.name = c("SCT.weight", "ADT.weight"))
  
  
  Tcell.seurat <- RunUMAP(Tcell.seurat, 
                          nn.name = "weighted.nn", 
                          reduction.name = "wnn.umap", 
                          reduction.key = "wnnUMAP_")
  
  Tcell.seurat <- FindClusters(Tcell.seurat, 
                               graph.name = "wsnn", 
                               algorithm = 3, 
                               resolution = 1.5, 
                               verbose = FALSE)
  
  
  DimPlot(Tcell.seurat,
          reduction = "wnn.umap", 
          label = T,
          group.by = "wsnn_res.1.5")
  
  
}
```



### Save Tcell only seurat object
```{r save_Tcell_only_seurat_object}

if(long.compute){
  
  
  SaveSeuratRds(Tcell.seurat, file = "saves/Preprocess_Tcell_seurat.rds")
  
}

```

